Exploiting the Graph: Increasing Engagement in Collaborative Interactive Evolution

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
A few corrections

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Interactive evolution is the leading technique employed by 
Evolutionary Art researchers, where users' preferences guide the search. A common
practice is to develop Interactive Evolutionary Computation (IEC) algorithms
as web applications in order to give users easy access the application as they
often depend on volunteers who visit the system to interactively help with the search. 
There are several drawbacks that limit user participation 
such as human fatigue and boredom resulting from evaluating 
a large number of individuals. 
To tackle this issues, we propose the design of IEC systems using a human-centered 
approach, with a framework consisting of a social network of 
volunteers, interacting with a population consisting also of a network 
of individuals. In this work we propose a graph model as a practical 
and efficient tool for mapping in a natural way the relationships 
found in the system. A case study is presented as a proof of concept, 
providing both conceptual and implementation details of a graph based model 
as it is  applied in the implementation of an IEA using the EvoSpace-Interactive platform. 
As a result, the graph model was successfully applied in a gamification technique 
developed to increase user engagement.

@@

The remainder of this paper is organized as follows.
Section \ref{sec:interactive} presents related work on the topic 
of Interactive Evolution with emphasis in data models.
Then, Section \ref{sec:evospace-i} presents the computational platform on which 
the main proposal of this work is developed. The proposed graph model is 
presented in Section \ref{sec:graph}.
The experimental work and results are presented in Section \ref{sec:experiments}.
Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

@@

The IEC the approach has been used in a wide variety of applications and problem domains
and remains an active area of research. However, in this work, emphasis is given to IEAs 
implemented as volunteer based collaborative systems; where volunteers interact and evaluate 
an evolving population, thus guiding the search based on an aggregate their subjective 
preferences. What follows is not intended as a comprehensive survey, only a review of 
the most relevant contributions to the present work, with special attention put in
data structures and data management requirements.

From the earlier works in the IEC field the notion of user collaboration in a natural way
has been proposed. A notable application was the Galapagos Project \cite{sims1997interactivity},
an exhibit in the Tokyo Multimedia Museum (1997--2000) were visitors interacted with images presented in 
twelve displays by selecting those they find most aesthetically interesting by standing on
step sensors in front of each. An early example of web based collaboration is the work by 
Langdon \cite{langdon:2004}, evolving fractal representations of virtual creatures. It proposed a distributed 
EA using a global population stored on a central web server sending 
portions of the population to remote clients using JavaScript. Users evaluated individuals locally 
and returned to the server those they liked more, to be distributed over the web.
Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} use web-based IEAs 
to evolve artistic artifacts using a generative encoding, compositional pattern producing networks.

In all these systems, the evolutionary interactive process is captured by the system. The genetic lineage of each
individual is presented where users can visualize the relationships between them and evolutionary operators used in
their creation. Also there is information regarding users and their collaborations.  Common offerings show 
interesting individuals order by categories, number of votes, best of the week among others. In a recent work by
Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
new designs of robot locomotiom. Collaboration is encoureged by gamifying the system, using the maximum distance indicator
to inspire the user to try and “beat” previous designs.

@@

Collaborative IEC systems need to store highly connected data, as it is common in current applications
like social networks. In order to deal with large datasets of connected data found in these systems,
graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
In this work the use of a graph data model for representing the interactive process of C-IEC systems is proposed.
Moreover, an proof-of-concept implemantation of IEC incorporating the graph model to the evolutionary algorithm is presented.

@@

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those individuals they like, or by taking individuals
from their friend\'s collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, an individual sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
He can assign a quality assessment to the individual using
a five star rating measure.
Additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special directory to store individuals a user prefers and wishes
to save. After the user finishes interacting the individual
on the Wall, he can choose to retrieve a new one from population.
At the left side, is the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, a user can browse the content of each collection and from
there share images through the social network.
When a user browses over an individual a detail pane shows how many users have
liked the individual. The pane also includes a link to the individual's 
details, the parents, genetic operators that created it, and genealogy information.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Interactive evolution is the leading technique employed by 
Evolutionary Art researchers, where users' preferences guide the search. A common
practice is to develop Interactive Evolutionary Computation (IEC) algorithms
as web applications in order to give users easy access the application as they
often depend on volunteers who visit the system to interactively help with fitness assignement. 
There are several drawbacks that limit user participation 
such as human fatigue and boredom resulting from evaluating 
a large number of individuals. 
To tackle these issues, we propose the design of IEC systems using a human-centered 
approach, with a framework consisting of a social network of 
volunteers, interacting with a population consisting also of a network 
of individuals. In this work we suggest the use a graph model as a practical 
and efficient tool for mapping the relationships 
found in the system. A case study is presented as proof-of-concept, 
providing both conceptual and implementation details of a graph based model 
as it is  applied in the implementation of an IEC system. 
As result show, the data model was successfully applied in a gamification technique 
developed to increase user engagement.

@@

The remainder of this paper is organized as follows.
Section \ref{sec:interactive} presents related work on the topic 
of Interactive Evolution.
Then, Section \ref{sec:evospace-i} presents the computational platform on which 
the main proposal of this work is developed. The proposed graph model is 
presented in Section \ref{sec:graph}.
The experimental work and results are presented in Section \ref{sec:experiments}.
Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

@@

The IEC the approach has been used in a wide variety of applications and problem domains
and remains an active area of research. However, in this work, emphasis is given to IEAs 
implemented as volunteer based collaborative systems; where volunteers interact and evaluate 
an evolving population, thus guiding the search based on an aggregate of their subjective 
preferences. What follows is not intended as a comprehensive survey, only a review of 
the most relevant contributions to the present work, with special attention put in
data structures and data management requirements.

From the earlier works in the IEC field the notion of user collaboration in a natural way
has been proposed. A notable application was the Galapagos Project \cite{sims1997interactivity},
an exhibit in the Tokyo Multimedia Museum (1997--2000) were visitors interacted with images presented in 
twelve displays by selecting those they find most aesthetically interesting by standing on
step sensors in front of them. An early example of web based collaboration is the work by 
Langdon \cite{langdon:2004}, evolving fractal representations of virtual creatures. It proposed a 
distributed EA using a global population stored on a central web server sending 
portions of the population to remote clients using JavaScript. Users evaluated individuals locally 
and returned to the server those they liked more, to be distributed over the web.
Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} use web-based IEAs 
to evolve artistic artifacts using a generative encoding, compositional pattern producing networks.

In all these systems, the interactive process is captured by the system. The genetic lineage of each
individual is presented in a graphical interface 
where users can visualize the relationships between them, and evolutionary operators used in
their creation; they also display information regarding users and their collaborations.  Common offerings show 
interesting individuals ordered by categories, number of votes, and best of the week among others. In a recent work by
Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
new designs of robot locomotiom. Collaboration is encoureged by gamifying the system 
using the maximum distance indicator to inspire the user to try and “beat” previous designs.

@@

Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common in current applications
like social networks. In order to deal with large datasets of connected data found in these systems,
graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
In this work the use of a graph data model for representing the interactive process of C-IEC systems is proposed.
Moreover, an proof-of-concept implemantation of an IEA incorporating the graph model is presented.

@@

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those individuals they like, or by taking individuals
from their friend\'s collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, an individual sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the individual using
a five star rating measure or
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder to store those individuals a user prefers and wishes
to save. After the user finishes interacting with the individual
on the Wall, he can choose to retrieve a new one from the population.
At the left side, is the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, a user can browse the content of each collection and from
there share images through the social network.
When a user browses over an individual a detail pane shows how many users have
liked the individual. The pane also includes a link to the individual's 
details, the parents, genetic operators that created it, and genealogy information.
